### 2024-12-02 manuscript.tex
Update on Overleaf.

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is the 5-km Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). The high resolution allows for parameterizations of convection and
clouds to be avoided. Standard-resolution models have substantial cloud biases
over the Southern Ocean (SO), affecting radiation and sea surface temperature.
We evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation levels being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Less stable conditions are the most problematic for
ICON and the reanalyses.  In daily cloud cover, ICON and the reanalyses tend to
be about 1 and 2 oktas clearer, respectively. Compared to radiosondes,
potential temperature is accurate in the reanalyses, but ICON underestimates
stability over the low-latitude SO and too humid in the boundary layer. MERRA-2
is too humid at all heights. SO cloud biases remain a substantial issue in the
GSRM, but are an improvement over the lower-resolution reanalyses. Explicitly
resolved convection and cloud processes were not enough to address the model
cloud biases.

@@

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \cite{schuddeboom2021} over the high-latitude SO, whereas
over the low-latitude SO it tends to be more neutral or negative in some
models. Too much absorbed SW radiation over the SO was also identified in the
GSRM SCREAM \cite{caldwell2021}. Compensating biases are possible, such as the
`too few too bright' cloud bias, characterised by too small cloud fraction and
too large cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a reasonable SW radiation balance by
reflecting too much SW radiation from clouds, but has too little cloud area
overall. A study by \cite{konsta2022} showed that this type of bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \cite{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to Moderate Resolution Imaging Spectroradiometer
(MODIS) data] and overestimates downwelling TOA SW radiation on by
approximately 10 Wm$^\mathrm{-2}$ [relative to the Clouds and the Earth’s
Radiant Energy System (CERES) data] in the highest model resolution run (2.5
km).  \cite{zhao2022} reported a similar SW radiation bias in five analysed
CMIP6 models over high-latitude SO and the total cloud fraction underestimation
on the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulations over a SO
region adjacent to Tasmania against the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations \cite{mcfarquhar2021}. They found
the ICON cloud optical thickness was underestimated relative to Himawari‐8
satellite observations, but also identified large differences in cloud top
phase.

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere, caused by too much shortwave heating of the surface, too
little longwave cooling of the surface, or in the ocean circulation.
Interactions of both are also possible, for example SST affecting clouds and
clouds affecting the surface radiation. \cite{zhang2023} has shown that SST
biases have improved in CMIP6 compared to CMIP5 [relative to the European
Centre for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5; ERA5], with SST
overall increasing in the later CMIP phase. However, over the SO this resulted
in an even higher positive bias, especially in the Atlantic Ocean (AO) sector
of the SO, increasing by up to 1°C.  \cite{luo2023} identified that the SO SST
bias in an ensemble of 18 CMIP6 models originates not from the surface heat and
radiation fluxes (using reanalyses as a reference), but from a warm bias in the
Northern Atlantic Deep Water.

The main aim of this study is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst, Max-Planck-Institute for
Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of
Technology, and the Center for Climate Systems Modeling.  Previous studies have
identified substantial large-scale biases in climate model clouds over the SO,
affecting sea surface temperature and the Earth’s albedo. Our aim is to
quantify how well the GSRM ICON is simulating clouds in this region,
particularly in light of the fact that subgrid-scale clouds and convection are
not parameterized in this model. This region is mostly dominated by boundary
layer clouds generated by shallow convection, and these are problematic to
observe by spaceborne lidars and radars, which are affected by attenuation by
overlapping and thick clouds and ground clutter, respectively.  Specifically,
the radar on CloudSat and lidar on CALIPSO (neither of which are now
operational) are affected by the abovementioned issues, resulting in a strong
underestimation of cloud occurrence below 2 km relative to ground-based lidar
observations \citep{mcerlich2021}.  This, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

@@

TOA radiation quantities are taken from the CERES instruments on board the
Terra and Aqua satellites \citep{wielicki1996,loeb2018}. In our analysis we
used the adjusted all sky SW and LW upwelling fluxes at TOA from the synoptic
TOA and surface fluxes and clouds 1 degree daily edition 4A product
(CER\_SYN1deg-Day\_Terra-Aqua-MODIS\_Edition4A)
\citep{doelling2013,doelling2016}.

Radiation calculations presented in the results (Section \ref{sec:results})
were done in such a way that they always represent averages of daily means.
This was done in order to be consistent with the CERES SYN1deg data, which are
available as daily means. Therefore, every instantaneous profile in the
simulated lidar data was assigned a daily mean radiation value corresponding to
the day (in the Coordinated Universal Time; UTC). In turn, the average
radiation during the entire voyage or station observation period were
calculated as the average of the profile values. In the observed lidar data,
the daily mean radiation value was taken from the spatially and temporally
co-located CERES SYN1deg data of the day (in UTC). The voyage or station
average was calculated in the same way.

@@

Precipitation can cause strong enough lidar backscattering to be recognised as
clouds by the threshold-based cloud detection method used in the ALCF. This is
undesirable if equivalent precipitation backscatter is not included in the
simulated lidar profiles. It was not possible to include precipitation
simulation in the ALCF due to the absence of required fields in the ICON model
output and the reanalysis data (the liquid and ice precipitation mass mixing
ratios). The required radiation calculations for precipitation are also
currently not implemented in the ALCF, even though this is a planned feature.
In order to achieve a fair comparison of observations with models, one should
exclude observed and simulated lidar profiles with precipitation either
manually or using an automated method. It is relatively difficult to
distinguish precipitation backscatter from cloud backscatter in lidar
observations, especially when only one wavelength channel and no polarised
channel are available. In models, the same can be accomplished relatively
easily by excluding profiles exceeding a certain amount of surface
precipitation flux. In the observations, using precipitation flux measurements
from rain gauges can be very unreliable on ships due to ship movement,
turbulence caused by nearby ship structures, and sea spray. Our analysis of
rain gauge data from the RV \emph{Tangaroa} showed large discrepancies between
the rain gauge time series and human-performed synoptic observations, as well
as large inconsistencies in the rain gauge time series. Human-performed
observations of precipitation presence or absence are expected to be reliable
but only cover a limited set of time instants. Therefore, it was desirable to
implement a method of detecting precipitation from observed backscatter
profiles alone.

@@

In addition to the above, we partitioned our data into two mutually exclusive
subsets by stability. We determined this by calculating lower tropospheric
stability (LTS) as the difference between the potential temperature at 700 hPa
and the surface.  Based on a histogram of LTS in ERA5 and MERRA-2 calculated at
all voyage tracks and stations (Fig.  \ref{fig:lts}), we determined a dividing
threshold of 12 K for relatively unstable (< 12 K) and relatively stable (>= 12
K) conditions.

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relevant for the subsequent analysis, because these
conditions are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a, b show a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied on the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, differences in the strength of the local
extremes of occurrence are notable, especially over the Amundsen Sea, which is
more cyclonic in the reanalysis, and around Cape Adare, which is more cyclonic
in ICON. These differences might, however, stem from the relatively short time
periods of comparison (4 years) and the fact that the model is free-running.

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON
and the reanalyses by about 10\% and 20\%, respectively. When analysed by
height, ICON overestimates cloud occurrence below 1 km and underestimates it
above, MERRA-2 underestimates cloud occurrence at all heights, especially near
the surface, and ERA5 simulates cloud occurrence relatively well above 1 km,
but strongly underestimates it near the surface.  We note that fog or
near-surface clouds are strongly lacking in the reanalyses (fog and clouds are
both included in the cloud occurrence).  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. We conclude that the ICON results are overall
better matching the observations than the reanalyses in this metric.

In the general case (Fig. \ref{fig:cloud-occurrence}a), the observations show
cloud occurrence peaking at the surface, whereas models show a higher peak (at
about 500 m). The models underestimate the total cloud fraction by 10--20\% and
show a strong drop in cloud occurrence near the surface, but this is not
supported by the observations. ICON and ERA5 overestimate cloud occurrence at
the peak (between 0--1 km). Above 1 km, ICON and MERRA-2 underestimate cloud
occurrence, but ERA5 is very accurate. The exaggerated peak in models is partly
supported by the lifting condensation level (LCL) distribution, which peaks
higher in the models than in the observations (at the surface), although this
is not very pronounced.

When subsetted by low- and high-latitude zones (Fig.
\ref{fig:cloud-occurrence}b, c), we see that the low-latitude SO zone shows
a stronger peak of cloud occurrence near the surface than the high-latitude SO
zone, and this could be because higher latitudes have more unstable atmospheric
profiles. The low- and high-latitude SO zones show similar biases in models as
in the general case, but ERA5 does not overestimate the peak in the
low-latitude SO zone (near-surface cloud occurrence is still strongly
underestimated).

When subsetted by cyclonic and non-cyclonic situations (Fig.
\ref{fig:cloud-occurrence}d, e), we see that the cyclonic situations have a
larger amount of observed cloudiness, including the peak and total cloud
fraction. In these situations, the models are doing a relatively good job of
getting the vertical profile of cloud occurrence right, but still tend to
underestimate cloud occurrence above 1 km and near the surface. Non-cyclonic
situations are similar to the general case.

When subsetted by relatively stable and unstable conditions (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in relatively stable
situations cloud occurrence peaks strongly at the surface in observations,
compared to relatively unstable situations, where the peak is more obtuse
and spread across the altitudes of 0--1 km.  In relatively stable
situations, the models are doing a fairly good job, but overestimate cloud
occurrence at the peak below 1 km; above 1 km, they show similar biases as in
the general case.  In relatively unstable situations, the bias in ICON is
very pronounced, with a much stronger peak at about 500 m, ERA5 is
underestimating cloud occurrence below 1 km (especially near the surface), and
MERRA-2 is underestimating it even more strongly.

In all situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in relatively unstable and
non-cyclonic conditions. It also shows a peak of cloud occurrence at higher
altitude than observations (500 m vs. near the surface), and correspondingly,
its LCL tends to be also higher. MERRA-2 can be generally characterised as
underestimating cloud occurrence at nearly all altitudes as well as the total
cloud fraction, but mostly above and below 500 m (the peak at 500 m is
well-represented). It struggles the most in the low-latitude SO zone and in the
relatively unstable situations. ERA5 can be generally characterised as
representing cloud occurrence correctly above about 1--1.5 km, overestimating
below, but underestimating near-surface cloud occurrence (0--500 m). The total
cloud fraction is strongly underestimated in all situations. It has a tendency
towards underestimation in the low-latitude SO zone and relatively unstable
situations; conversely, overestimating in the high-latitude SO zone and the
relatively stable conditions.

@@

We analysed the daily cloud cover (total cloud fraction) distribution. This is
a measure of cloudiness, irrespective of height, calculated over the course of
a day (UTC). A cloud detected at any height means that the lidar profile was
classified as cloudy; otherwise, it was classified as a clear sky. When all
profiles in a day are taken together, the cloud cover for the day is defined as
the fraction of cloudy profiles in the total number of profiles, expressed in
oktas (multiples of 1/8).

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets of data
as in the previous section. Observations have the greatest representation of
high cloud cover (5--8 oktas), peaking at 7 oktas. This is not represented by
ICON or the reanalyses.  While ICON is the closest, it tends to be 1 okta
clearer than the observations, peaking at 6 oktas, and highly underestimating
days with 8 oktas.  Overall, the reanalyses show results similar to each other,
underestimating cloud cover by about 2 oktas and strongly underestimating days
with 7 and 8 oktas. Of the two reanalyses, MERRA-2 shows slightly higher cloud
cover and thus is more consistent with observations.

When analysed by subsets, observations in the cyclonic subset show the highest
cloud cover, with 8 oktas occurring on one half of such days (Fig.
\ref{fig:cloud-cover}d).  This is not represented by ICON or the reanalyses at
all. Interestingly, clear sky days (0 oktas) also have a local maximum peaking
at about 15\% in this subset.  When we contrast the low- and high-latitude
zones, we see that the high-latitude zone tends to have greater cloud cover,
peaking at 8 oktas (Fig.  \ref{fig:cloud-cover}c). The high-latitude zone also
has almost no clear sky or small cloud cover cases (0--4 oktas). ICON and the
reanalyses represent at least this characteristic of the distribution well for
0--3 oktas, but otherwise show biases similar to the general case. One of the
greatest biases is present in ERA5 in the relatively unstable subset, in which
ERA5 peaks at 3 oktas, whereas the observations peak at 7 oktas and show
negligible cloud cover below 5 oktas.

@@

Overall, the mean $\theta_v$ is relatively well represented in ICON and
MERRA-2, being only slightly colder in the mid-to-high troposphere (less
stable) in ICON than in observations (Fig. \ref{fig:potential-temperature}a).
Large differences exist, however, in the 40--55°S zone, where ICON is colder in
$\theta_v$ by up to about 5 K and more so at higher altitudes (Fig.
\ref{fig:potential-temperature}b). In other subsets, the bias is relatively
small. MERRA-2 is very close to the observations, possibly due to a high
accuracy of assimilation of this quantity. Notably, the variability of virtual
potential temperature (as represented by the percentiles) is much smaller in
ICON than in the observations. This indicates that the model's internal
variability in the lower-tropospheric thermodynamic conditions in the SO is
smaller than in reality.

Relative humidity displays much larger biases. In all subsets, ICON is too
humid in the first 1 km but very accurate above, except for the 40--55°S zone
and unstable conditions (Fig. \ref{fig:potential-temperature}b, g),  where it
is too dry between about 1 and 3 km. MERRA-2, on the other hand, is much more
humid than observations at all altitudes and in all subsets, by up to about
20\% at 5 km.

\section{Limitations}

>>>>

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). The high resolution allows for parameterizations of convection and
clouds to be avoided. Standard-resolution models have substantial cloud biases
over the Southern Ocean (SO), affecting radiation and sea surface temperature.
We evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation levels being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Less stable conditions are the most problematic for
ICON and the reanalyses.  In daily cloud cover, ICON and the reanalyses tend to
be about 1 and 2 oktas clearer, respectively. Compared to radiosondes,
potential temperature is accurate in the reanalyses, but ICON underestimates
stability over the low-latitude SO and too humid in the boundary layer. MERRA-2
is too humid at all heights. SO cloud biases remain a substantial issue in the
GSRM, but are an improvement over the lower-resolution reanalyses. Explicitly
resolved convection and cloud processes were not enough to address the model
cloud biases.

@@

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \cite{schuddeboom2021} over the high-latitude SO, whereas
over the low-latitude SO it tends to be more neutral or negative in some
models. Too much absorbed SW radiation over the SO was also identified in the
GSRM SCREAM \cite{caldwell2021}. Compensating biases are possible, such as the
`too few too bright' cloud bias, characterised by too small cloud fraction and
too large cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a reasonable SW radiation balance by
reflecting too much SW radiation from clouds, but has too little cloudarea
overall. A study by \cite{konsta2022} showed that this type of bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \cite{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to Moderate Resolution Imaging Spectroradiometer
(MODIS) data] and overestimates downwelling TOA SW radiation by
approximately 10 Wm$^\mathrm{-2}$ [relative to the Clouds and the Earth’s
Radiant Energy System (CERES) data] in the highest model resolution run (2.5
km).  \cite{zhao2022} reported a similar SW radiation bias in five analysed
CMIP6 models over high-latitude SO and the total cloud fraction underestimation
on the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulations over a SO
region adjacent to Tasmania against the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations \cite{mcfarquhar2021}. They found
the ICON cloud optical thickness was underestimated relative to Himawari‐8
satellite observations, but also identified large differences in cloud top
phase.

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere, caused by too much shortwave heating of the surface, too
little longwave cooling of the surface, or in the ocean circulation.
Interactions of both are also possible, for example SST affecting clouds and
clouds affecting the surface radiation. \cite{zhang2023} have shown that SST
biases have improved in CMIP6 compared to CMIP5 [relative to the European
Centre for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5; ERA5], with SST
overall increasing in the later CMIP phase. However, over the SO this resulted
in an even higher positive bias, especially in the Atlantic Ocean (AO) sector
of the SO, increasing by up to 1°C.  \cite{luo2023} identified that the SO SST
bias in an ensemble of 18 CMIP6 models originates not from the surface heat and
radiation fluxes (using reanalyses as a reference), but from a warm bias in the
Northern Atlantic Deep Water.

The main aim of this study is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst, Max-Planck-Institute for
Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of
Technology, and the Center for Climate Systems Modeling.  Previous studies have
identified substantial large-scale biases in climate model clouds over the SO,
affecting sea surface temperature and the Earth’s albedo. Our aim is to
quantify how well the GSRM ICON is simulating clouds in this region,
particularly in light of the fact that subgrid-scale clouds and convection are
not parameterized in this model. This region is mostly dominated by boundary
layer clouds generated by shallow convection, and these are problematic to
observe by spaceborne lidars and radars, which are affected by attenuation by
overlapping and thick clouds and ground clutter, respectively.  Specifically,
the radar on CloudSat and lidar on CALIPSO (neither of which are nowoperational) are affected by the abovementioned issues, resulting in a strong
underestimation of cloud occurrence below 2 km relative to ground-based lidar
observations \citep{mcerlich2021}.  This, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

@@

TOA radiation quantities are taken from the CERES instruments onboard the
Terra and Aqua satellites \citep{wielicki1996,loeb2018}. In our analysis we
used the adjusted all sky SW and LW upwelling fluxes at TOA from the synoptic
TOA and surface fluxes and clouds 1 degree daily edition 4A product
(CER\_SYN1deg-Day\_Terra-Aqua-MODIS\_Edition4A)
\citep{doelling2013,doelling2016}.

Radiation calculations presented in the results (Section \ref{sec:results})
were completed such that they always represent daily means in order to be consistent with the CERES SYN1deg data. Therefore, every instantaneous profile in the
simulated lidar data was assigned a daily mean radiation value corresponding to
the day (in the Coordinated Universal Time; UTC). In turn, the average
radiation during the entire voyage or station observation period were
calculated as the average of the profile values. In the observed lidar data,
the daily mean radiation value was taken from the spatially and temporally
co-located CERES SYN1deg data for the day (in UTC). The voyage or station
average was calculated in the same way.

@@

Precipitation can cause strong enough lidar backscattering to be recognised as
clouds by the threshold-based cloud detection method used in the ALCF. This is
undesirable if equivalent precipitation backscatter is not included in the
simulated lidar profiles. It was not possible to include precipitation
simulation in the ALCF due to the absence of required fields in the ICON model
output and the reanalysis data (the liquid and ice precipitation mass mixing
ratios). The required radiation calculations for precipitation are also
currently not implemented in the ALCF, even though this is a planned future addition.
In order to achieve a fair comparison of observations with model output, we
exclude observed and simulated lidar profiles with precipitation either
manually or using an automated method. It is relatively difficult to
distinguish precipitation backscatter from cloud backscatter in lidar
observations, especially when only one wavelength channel and no polarised
channel are available. In models, the same can be accomplished relatively
easily by excluding profiles exceeding a certain surface
precipitation flux. In the observations, using precipitation flux measurements
from rain gauges can be very unreliable on ships due to ship movement,
turbulence caused by nearby ship structures, and sea spray. Our analysis of
rain gauge data from the RV \emph{Tangaroa} showed large discrepancies between
the rain gauge time series and human-performed synoptic observations, as well
as large inconsistencies in the rain gauge time series. Human-performed
observations of precipitation presence or absence are expected to be reliable
but only cover a limited set of times. Therefore, it was desirable to
implement a method of detecting precipitation from observed backscatter
profiles alone.

@@

In addition to the above, we partitioned our data into two mutually exclusive
subsets based on lower tropospheric stability (LTS), which is derived as the difference between the potential temperature at 700 hPa
and the surface.  Based on a histogram of LTS in ERA5 and MERRA-2 calculated at
all voyage tracks and stations (Fig.  \ref{fig:lts}), we determined a dividing
threshold of 12 K for relatively unstable (< 12 K) and relatively stable (>= 12
K) conditions.

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relevant for the subsequent analysis, because these
conditions are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a, b show a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied to the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, differences in the strength of the local
extremes of occurrence are notable, especially over the Amundsen Sea, which is
more cyclonic in the reanalysis, and around Cape Adare, which is more cyclonic
in ICON. These differences might, however, stem from the relatively short time
periods of comparison (4 years) and the fact that the model is free-running.

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON
by about 10\% and in the reanalyses by about 20\%. When analysed by
height, ICON overestimates cloud occurrence below 1 km and underestimates it
above, MERRA-2 underestimates cloud occurrence at all heights, especially near
the surface, and ERA5 simulates cloud occurrence relatively well above 1 km,
but strongly underestimates it near the surface.  We note that fog or
near-surface clouds are strongly underestimated in the reanalyses (fog and clouds are
both included in the cloud occurrence).  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. We conclude that the ICON results ch the observations better than the reanalyses in this metric.

For all observations considered (Fig. \ref{fig:cloud-occurrence}a), the data show
cloud occurrence peaking at the surface, whereas the models show a higher peak (at
about 500 m). The models generally underestimate the total cloud fraction by 10--20\% and show a strong reduction in cloud occurrence near the surface, which is not
supported by the observations. ICON and ERA5 overestimate cloud occurrence at
their peak (between 0--1 km). Above 1 km, ICON and MERRA-2 underestimate cloud
occurrence, but ERA5 is very accurate. The exaggerated peak in models is partly
supported by the lifting condensation level (LCL) distribution, which peaks
higher in the models than in the observations (at the surface), although this
is not very pronounced.

When subsetted by low- and high-latitude SO zones (Fig.
\ref{fig:cloud-occurrence}b, c), we see that the low-latitude SO zone displays
a stronger peak of cloud occurrence near the surface than the high-latitude SO
zone, and this could be because higher latitudes have more unstable atmospheric
profiles. The low- and high-latitude SO zones show similar biases in models as
in the general case, but ERA5 does not overestimate the peak in the
low-latitude SO zone (near-surface cloud occurrence is still strongly
underestimated).

When subsetted by cyclonic and non-cyclonic situations (Fig.
\ref{fig:cloud-occurrence}d, e), we see that the cyclonic situations have a
larger amount of observed cloudiness, including peak and total cloud
fraction. In these situations, the model vertical profiles of cloud occurrence compare well with observations, but still tend to
underestimate cloud occurrence above 1 km and near the surface. Non-cyclonic
situations are similar to the general case.

When subsetted by relatively stable and unstable conditions (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in relatively stable
situations cloud occurrence peaks strongly at the surface in observations,
compared to relatively unstable situations, where the peak is more obtuse
and spread across the altitudes of 0--1 km.  In relatively stable
situations, the models do a fairly good job, but overestimate cloud
occurrence at the peak below 1 km; above 1 km, they show similar biases as in
the general case.  In relatively unstable situations, the bias in ICON is
very pronounced, with a much stronger peak at about 500 m, ERA5 underestimates cloud occurrence below 1 km (especially near the surface), and
MERRA-2 underestimates cloud occurrence even more strongly.

In all situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in relatively unstable and
non-cyclonic conditions. It also shows a peak of cloud occurrence at higher
altitude than observations (500 m vs. near the surface), and correspondingly,
its LCL tends to be higher. MERRA-2 can be generally characterised as
underestimating cloud occurrence at nearly all altitudes as well as the total
cloud fraction, but mostly above and below 500 m (the peak at 500 m is
well-represented). MERRA-2 displays the largest errors relative to observations in the low-latitude SO zone and in relatively unstable situations. ERA5 can be generally characterised as
representing cloud occurrence correctly above about 1--1.5 km, overestimating
below, but underestimating near-surface cloud occurrence (0--500 m). The total
cloud fraction is strongly underestimated in all situations. ERA5 has a tendency
towards underestimation in the low-latitude SO zone and relatively unstable
situations; conversely, overestimating in the high-latitude SO zone and relatively stable conditions.

@@

We also analysed the daily cloud cover (total cloud fraction) distribution. This is
a measure of cloudiness, irrespective of height, calculated over the course of
a day (UTC). A cloud detected at any height means that the lidar profile was
classified as cloudy; otherwise, it was classified as clear sky. When all
profiles in a day are taken together, the cloud cover for the day is defined as
the fraction of cloudy profiles in the total number of profiles, expressed in
oktas (multiples of 1/8).

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets of data
as in Section \ref{sec:cloud-occurrence}. Observations display the highest proportion of
high cloud cover values (5--8 oktas), peaking at 7 oktas. This pattern is not represented by
ICON or either reanalyses.  While ICON is closest to matching the observed distribution, it tends to be 1 okta
clearer than the observations, peaking at 6 oktas, and highly underestimating
days with 8 oktas.  Overall, the reanalyses show results similar to each other,
underestimating cloud cover by about 2 oktas and strongly underestimating days
with 7 and 8 oktas. Of the two reanalyses, MERRA-2 shows slightly higher cloud
cover and thus is more consistent with observations.

When analysed by subsets, observations in the cyclonic subset show the highest
cloud cover, with 8 oktas occurring on one half of such days (Fig.
\ref{fig:cloud-cover}d).  This sensitivity to cyclonic conditions is not observed in ICON or the reanalyses. Interestingly, clear sky days (0 oktas) also have a local maximum peaking
at about 15\% in this subset.  When we contrast the low- and high-latitude
zones, we see that the high-latitude zone tends to have greater cloud cover,
peaking at 8 oktas (Fig.  \ref{fig:cloud-cover}c). The high-latitude zone also
has almost no clear sky or small cloud cover cases (0--4 oktas). ICON and the
reanalyses represent this characteristic of the distribution well for
0--3 oktas, but otherwise show biases similar to the general case. One of the
greatest biases is present in ERA5 in the relatively unstable subset, in which
ERA5 peaks at 3 oktas, while the observations peak at 7 oktas and show
negligible cloud cover below 5 oktas.

@@

Overall, the mean $\theta_v$ is very well represented in ICON and
MERRA-2, being only slightly colder in the mid-to-high troposphere (less
stable) in ICON than in observations (Fig. \ref{fig:potential-temperature}a).
Larger differences exist, however, in the 40--55°S zone, where ICON is colder by about 5 K at higher altitudes (Fig.
\ref{fig:potential-temperature}b). In other subsets, the bias is relatively
small. MERRA-2 is very close to the observations, possibly due to a high
accuracy of assimilation of this quantity. Notably, the variability of virtual
potential temperature (as represented by the percentiles) is much smaller in
ICON than in the observations. This indicates that the model's internal
variability in the lower-tropospheric thermodynamic conditions in the SO is
smaller than in reality.

Relative humidity displays much larger biases. In all subsets, ICON is too
humid in the first 1 km by about 5\% but very accurate above, except for the 40--55°S zone
and unstable conditions (Fig. \ref{fig:potential-temperature}b, g),  where it
is too dry between about 1 and 3 km. MERRA-2, on the other hand, is much more
humid than observations at all altitudes and in all subsets, by up to about
20\% at 5 km.

\section{Limitations of this study}
